---
title: "DATA 607 Week 1 Assignment"
author: "Steve Tipton"
date: "February 4, 2018"
output: html_document
---

## Tasks

1. Import mushrooms data set
2. Create a data set with a subset of the columns (include edible/poisonous column and 3 or 4 others)
3. Add meaningful column names
4. Replace abbreviations in data

### Task 1 - Import data set

```{r setup}
mushroom <- read.csv("https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data", header = FALSE, sep = ",", stringsAsFactors = FALSE)
head(mushroom)
```

### Task 2 - Create subset

```{r}
mushSub <- subset(mushroom, select = c(1, 4, 6, 14, 15, 21, 22, 23))
head(mushSub)
```

### Task 3 - Add column names

```{r}
colnames(mushSub) <- c("class", "capColor", "odor", "stalkSurfaceBelowRing", "stalkColorAboveRing", "sporePrintColor", "population", "habitat")
head(mushSub)
```

### Task 4 - Replace abbreviations

```{r}
mushSub$class[mushSub$class == "p"] <- "poisonous"
mushSub$class[mushSub$class == "e"] <- "edible"
mushSub$capColor[mushSub$capColor == "n"] <- "brown"
mushSub$capColor[mushSub$capColor == "b"] <- "buff"
mushSub$capColor[mushSub$capColor == "c"] <- "cinnamon"
mushSub$capColor[mushSub$capColor == "g"] <- "gray"
mushSub$capColor[mushSub$capColor == "r"] <- "green"
mushSub$capColor[mushSub$capColor == "p"] <- "pink"
mushSub$capColor[mushSub$capColor == "u"] <- "purple"
mushSub$capColor[mushSub$capColor == "e"] <- "red"
mushSub$capColor[mushSub$capColor == "w"] <- "white"
mushSub$capColor[mushSub$capColor == "y"] <- "yellow"
mushSub$odor[mushSub$odor == "a"] <- "almond"
mushSub$odor[mushSub$odor == "l"] <- "anise"
mushSub$odor[mushSub$odor == "c"] <- "creosote"
mushSub$odor[mushSub$odor == "y"] <- "fishy"
mushSub$odor[mushSub$odor == "f"] <- "foul"
mushSub$odor[mushSub$odor == "m"] <- "musty"
mushSub$odor[mushSub$odor == "n"] <- "none"
mushSub$odor[mushSub$odor == "p"] <- "pungent"
mushSub$odor[mushSub$odor == "s"] <- "spicy"
mushSub$stalkSurfaceBelowRing[mushSub$stalkSurfaceBelowRing == "f"] <- "fibrous"
mushSub$stalkSurfaceBelowRing[mushSub$stalkSurfaceBelowRing == "y"] <- "scaly"
mushSub$stalkSurfaceBelowRing[mushSub$stalkSurfaceBelowRing == "k"] <- "silky"
mushSub$stalkSurfaceBelowRing[mushSub$stalkSurfaceBelowRing == "s"] <- "smooth"
mushSub$stalkColorAboveRing[mushSub$stalkColorAboveRing == "n"] <- "brown"
mushSub$stalkColorAboveRing[mushSub$stalkColorAboveRing == "b"] <- "buff"
mushSub$stalkColorAboveRing[mushSub$stalkColorAboveRing == "c"] <- "cinnamon"
mushSub$stalkColorAboveRing[mushSub$stalkColorAboveRing == "g"] <- "gray"
mushSub$stalkColorAboveRing[mushSub$stalkColorAboveRing == "r"] <- "green"
mushSub$stalkColorAboveRing[mushSub$stalkColorAboveRing == "p"] <- "pink"
mushSub$stalkColorAboveRing[mushSub$stalkColorAboveRing == "u"] <- "purple"
mushSub$stalkColorAboveRing[mushSub$stalkColorAboveRing == "e"] <- "red"
mushSub$stalkColorAboveRing[mushSub$stalkColorAboveRing == "w"] <- "white"
mushSub$stalkColorAboveRing[mushSub$stalkColorAboveRing == "y"] <- "yellow"
mushSub$sporePrintColor[mushSub$sporePrintColor == "k"] <- "black"
mushSub$sporePrintColor[mushSub$sporePrintColor == "n"] <- "brown"
mushSub$sporePrintColor[mushSub$sporePrintColor == "b"] <- "buff"
mushSub$sporePrintColor[mushSub$sporePrintColor == "h"] <- "chocolate"
mushSub$sporePrintColor[mushSub$sporePrintColor == "r"] <- "green"
mushSub$sporePrintColor[mushSub$sporePrintColor == "o"] <- "orange"
mushSub$sporePrintColor[mushSub$sporePrintColor == "u"] <- "purple"
mushSub$sporePrintColor[mushSub$sporePrintColor == "w"] <- "white"
mushSub$sporePrintColor[mushSub$sporePrintColor == "y"] <- "yellow"
mushSub$population[mushSub$population == "a"] <- "abundant"
mushSub$population[mushSub$population == "c"] <- "clustered"
mushSub$population[mushSub$population == "n"] <- "numerous"
mushSub$population[mushSub$population == "s"] <- "scattered"
mushSub$population[mushSub$population == "v"] <- "several"
mushSub$population[mushSub$population == "y"] <- "solitary"
mushSub$habitat[mushSub$habitat == "g"] <- "grasses"
mushSub$habitat[mushSub$habitat == "l"] <- "leaves"
mushSub$habitat[mushSub$habitat == "m"] <- "meadows"
mushSub$habitat[mushSub$habitat == "p"] <- "paths"
mushSub$habitat[mushSub$habitat == "u"] <- "urban"
mushSub$habitat[mushSub$habitat == "w"] <- "waste"
mushSub$habitat[mushSub$habitat == "d"] <- "woods"

head(mushSub)
```

## Further questions

I chose the variables based on the data dictionary's rules for identifying poisonous mushrooms.  I'll further subset the data set to explore those conditions and find the edible mushrooms using the negations of those rules.

### The poisonous rules (disjuntive)

1. odor=NOT(almond.OR.anise.OR.none)

```{r}
p1 <- subset(mushSub, !(odor %in% c("almond","anise","none")))
table(p1$odor, p1$class)
```

2. spore-print-color=green

```{r}
p2 <- subset(mushSub, sporePrintColor == "green")
table(p2$sporePrintColor, p2$class)
```

3. odor=none.AND.stalk-surface-below-ring=scaly.AND.(stalk-color-above-ring=NOT.brown)

```{r}
p3 <- subset(mushSub, odor == "none" & stalkSurfaceBelowRing == "scaly" & stalkColorAboveRing != "brown")
table(p3$stalkColorAboveRing, p3$class)
```

4. habitat=leaves.AND.cap-color=white

```{r}
p4 <- subset(mushSub, habitat == "leaves" & capColor == "white")
table(p4$habitat, p4$capColor, p4$class)
```

4a. population=clustered.AND.cap_color=white

```{r}
p4a <- subset(mushSub, population == "clustered" & capColor == "white")
table(p4a$population, p4a$capColor, p4a$class)
```

### Finding the edible mushrooms

By taking the negation of all the poisonous rules together, I can discover the edible mushrooms within the data set.  I can also identify how many poisonous mushrooms come up in this subset and calculate the accuracy of the rules.  Note that rule p3 is removed from the logical statement as it conflicts with rule p1.

1. Using rule p4:

```{r}
e1 <- subset(mushSub, 
             odor %in% c("almond", "anise", "none") & #p1
             sporePrintColor != "green" & #p2
             #(odor != "none" | stalkSurfaceBelowRing != "scaly" | stalkColorAboveRing == "brown") & #p3
             (habitat != "leaves" | capColor != "white")) #p4
table(e1$class)
```

With 40 poisonous mushrooms in the subset, the rules are 99.51% accurate when using p4.

2. Using rule p4a

```{r}
e1 <- subset(mushSub, 
             odor %in% c("almond", "anise", "none") & #p1
             sporePrintColor != "green" & #p2
             #(odor != "none" | stalkSurfaceBelowRing != "scaly" | stalkColorAboveRing == "brown") & #p3
             (population != "clustered" | capColor != "white")) #p4a
table(e1$class)
```

Substituting rule p4a for p4 also results in 40 poisonous mushrooms and the same accuracy rate of 99.51%